How to Learn Useful Changes to Software Projects\\(to Reduce Runtimes and Software Defects)

### 2015-08-26 Manuscript.tex
changing naveen's location
Updates from ShareLaTeX

\newcommand{\quart}[4]{\begin{picture}(100,4)%1
{\color{black}\put(#3,2){\circle*{4}}\put(#1,2){\line(1,0){#2}}}\end{picture}}

@@

\title{How to Learn Useful Changes to Software Projects\\(to Reduce Runtimes and Software Defects)}
\numberofauthors{4} 
\author{  
\alignauthor
Rahul~Krishna\\Tim Menzies, Xipeng~Shen \\
       \affaddr{CS, NC State, USA}\\
       {\{i.m.ralk,~tim.menzies,\\xipengshen\}@gmail.com}
\alignauthor
Andrian Marcus \\
       \affaddr{CS, UT Dallas  } \\ 
       \affaddr{Texas, USA}\\ 
       {amarcus@utdallas.edu}
       \alignauthor
Naveen  \\ Lekkalapudi\\
 \affaddr{Bellhops } \\ 
       \affaddr{Washington D.C.}\\ 
       {navek91@gmail.com}
\alignauthor
Lucas Layman \\
       \affaddr{Fraunhofer CESE  } \\ 
       \affaddr{College Park, USA}\\ 
       {llayman@fc-md.umd.edu}
\setlength{\columnsep}{7mm}
}
\maketitle
\begin{abstract}
 Business users now demand more insightful
 analytics. Instead of just predicting 
 some result, they also need tools that generate ``plans'';
 i.e. specific suggestions on  what to change  in order to
 improve on the predicted values. Hence, we propose the  XTREE planner and  show that
 it runs better than   three alternative methods. For this paper's    case studies,
 XTREE reduced
 the expected values of defects \& runtimes of some software projects to    
 45\% \& 61.5\%  (median) and  22\% \& 6\% (best case) of the initial values, respectively.
\end{abstract}
\section{Introduction}
Business users are demanding  tools that support    business-level interpretations of their data. At a panel on software analytics at ICSE`12, industrial practitioners lamented the state of the art in software analytics~\cite{menzies12a}. Panelists commented  ``prediction is all well and good, but what about decision making?''. Note that these panelists were more interested in the interpretations and follow-up that occurs after the mining, rather than just the mining itself. So:
\bi
\item
Instead of just accepting {\em predictions} on how many software defects to expect, business users might now demand a {\em plan} to reduce the likelihood of those defects.
\item
Instead of just accepting {\em predictions} on the runtime time of their software, business users might now demand a {\em plan} to reduce that runtime. 
\ei
In response to this business-level demands for planners,
we propose a novel {\em planning} method called XTREE for learning changes to a software system
such that its performance ``improves'', according to some measure. This paper uses XTREE
to reduce  the expected value of the
defects     in Jureczko et al.'s    JAVA systems~\cite{jureczko10};
and the runtimes   in   software    configured by  Siegmund et al.~\cite{sven12}.

The contributions of this paper are (1)  
    the new XTREE    algorithm and (2) an evaluation strategy for   planners.
In the case studies of this paper  our evaluations shows
 XTREE performing significantly ``better'' than    planners  proposed in our prior   work~\cite{me12c,krishna15},
 where ``better'' means:
\bi
\item {\em Effective:}    i.e.  if  a plan is   applied  then  some  statistically significant  change  in expected values of the results will  be  observed. 
\item {\em Succinct:} i.e. given two plans with the same effectiveness, we prefer the shorter one.
Such shorter plans are easier to explain (less to show) and faster to implement (less to change).
\item {\em Surprising}; i.e. we want the planners to (occasionally) tell us things we do not know already (otherwise,
there would be no point to using the planner).
\ei 
The rest of this paper is describes our data, our planners, and the experiment that
ranks XTREE against alternate approaches.  This is followed by notes on related work
and validity.
Note that, to allow for reproducibility, all scripts and data used in this 
study are available on-line at github.com/ai-se/XTREE.

@@

Our planners   use tables of data with independent features and one dependent feature, called the 
``class''. Classes are weighted such that they indicate  what rows are ``good'' and what rows are ``bad''. Plans  change   a row such that it is more like they 
will be more similar to ``good'' than ``bad''.
Specifically, for every test example $Z$,   planners proposes a  plan $\Delta$ to
  adjust   feature $Z_j$:
\[
\forall \delta_j \in \Delta :  Z_j = Z_j + \delta_j
\]
For example, to simplify a  large bug-prone  method, our planners might suggest
to a developer to reduce its size (i.e.  refactor that code by, say, splitting it across
two simpler functions).

@@

This paper is about the next step {\em after} prediction. Suppose
a business user is presented a prediction and they do not like what they see; e.g. the runtimes are too long of the number of defects is too high. This user may
then ask a {\em planning}  question; i.e. ``what can we change to do better than that?''.

@@

To answer this question, we note
that data miners explore two  ``clouds'' of data: (1) the cloud of training examples and (2) the  cloud   of test examples.
For a visualization of these clouds, see \fig{howxy}.

\begin{figure}[!t]
  \includegraphics[width=\linewidth]{figs/twodee.eps} 
 % \includegraphics[width=1\linewidth]{figs/BDBC.eps}
\caption{Gray, red, green show (1) training examples, (2) test examples and 
  (3) tests that have been altered by planners.
  This figure uses axes generated from the first two components of a PCA analysis of all points. 
}\label{fig:howxy}
\end{figure}

We should mistrust the predictions made by a  model   if it is being applied to examples  that are
too far away from the
training cloud.
To test for ``too far'', we can run a data mining experiment that tests how well
a model learned from the training data applies to the test data. Such experiments return some performance value.

@@

That said, the above comes with some important caveats:
\bi
\item 
We   strongly recommend that predictors are assessed prior to planning. That
issue is explored further in the \tion{tesd}.
\item
Planners should be designed to be {\em trust increasing}. We list four such planning methods in \tion{planners}.
\item
Where possible, planners should be assessed via some external
oracle that can accurately assess new examples. For an example of that kind of analysis,
see  \tion{valid}.
\ei

@@

\caption{Siegmund data.
For SQLite, the data  contains 4,553 configurations for prediction modeling and 100 additional random configurations for prediction evaluation, see \cite{vapp}.}\label{fig:cpm}
\end{figure}

@@

In his textbook on empirical methods, Cohen~\cite{cohen95} advises that any supposedly better
algorithm should be baselined against simpler alternatives. Accordingly,
this section described XTREE (which we call Method4) against three  
alternatives.
  XTREE  uses the decision tree learner of \fig{where}.D.  It is new to this paper.
  The other methods use the   top-down
	bi-clustering method described in \fig{where}.C  which recursively divides the
	data in two  using a dimension that captures the greatest variability in the data. 
	We proposed Method 1 and 2   in 2012~\cite{me12c} while Methods 3 comes from research conducted earlier this year~\cite{krishna15}.

@@

Method3 connects  each centroid to a nearest neighbor
by {\em gradient}.
Each gradient has a (bottom,top) end labelled  ($C_i,C_j$) containing the  (worst,best) centroid performance scores, respectively..  
For each test instance, Method3 
find the nearest gradient, the runs up to the top  best end $C_j$, then extracts $B_j$ (which is the
 best-in-cluster associated with  $C_j$).
The returned plan is then computed  from the delta between the test case
and $B_j$.

@@

\section{Experiments}

@@

\fig{jur} and \fig{conf1} show the effectiveness of our methods seen in 40 repeats with each data set.
In these experiments,   the dependent variables of Jureczko and Siegmund data sets are discrete and continuous in nature, respectively. Hence, while choosing the predictor, we used Random Forest (1) as a classifier for Jureczko data and (2) as a regressor for Siegmund data.
\bi
\item For Siegmund data, we randomized the order of the data, training on one half while identifying treatment plans on the remaining test data. 
\item For the Jureczko data, we used the training and testing sets of \fig{j}. For these datasets,
 all the SMOTE-ing and Random Forst tunings (discussed in \tion{tesd})
occurred in the {\em train} phase of \fig{design}.
\ei
 \subsection{Experimental Results}

@@

% \begin{figure}[t]
% \begin{minipage}{0.5\textwidth}
% \includegraphics[width=0.9\linewidth]{figs/Deltas-Lucene.eps}
% \label{fig:delta_ivy}
% \end{minipage}
% \begin{minipage}{0.5\textwidth}
% \includegraphics[width=0.9\linewidth]{figs/Deltas-BDBJ.eps}
% \label{fig:delta_apache}
% \end{minipage}
% \caption{Performance Comparison showing the number of times a certain feature requires to be changed. X-axis lists the various features that make up a data set. The y-axis contains the frequency of change in \%.}\label{fig:changed}
% \end{figure}

@@

\begin{figure*}[!t]
\centering
\includegraphics[width=\linewidth]{figs/Deltas-both.eps}
\caption{Percent frequency for how often certain feature was changed bu a plan.}\label{fig:changed}
\end{figure*}

@@

The only coupling issue that XTREE   usually adds to its plans is {\em cbo} (which appears 55\% of the time
in \fig{changed}). But note  that this is {\em object} coupling measure, not class coupling. So here
XTREE is warning against, say, some factory class generating a large
community of agents, all of the same class, who
co-ordinate on some task. This is a different issue to the class redesign issue that would
be triggered by altering {\em loc}}.

@@

\subsection{  Evaluation Bias} Another threat to validity of this work is our use
of predictors learned on the training data to assess the impact of our planners.
This issues was discussed in detail in \tion{trust}.

@@

@@

One problem with these   SBSE methods is that they can  make extensive mutations to the data they are exploring. In the language
of \tion{trust}, these methods may not be {\em trust-increasing} those algorithms make no attempt
to mutate new examples away from the kinds of data used to commission the model (in which case, we would
start doubting the model's output).

@@

Four planners were assessed here for the tasks of reducing defects and runtimes. 
Three of the methods come from our prior publications~\cite{me12c,krishna15}, and the conclusion of this
paper is that a novel fourth method clearly out-performs the other three
(measured in terms of {\em effectiveness, succinctness, and surprise}).
The reason for the superior performance of our new method is clear: when planning how to change
to different class methods, it is best to use a learner that reflects over those classes
as it divides the data.

Of course, there are many more methods for generating plans and
no   one paper can survey them all.
But the goal of this paper is not to claim that (e.g.) XTREE is some absolute optimal algorithm. Rather, it is
to offer a baseline result (with XTREE) and an  evaluation strategy that  can assess  if alternate methods are better than XTREE.
Our hope is that other   researchers apply this strategy  to repeat and/or improve
and/or refute our results.

>>>>

\newcommand{\quart}[4]{\begin{picture}(80,4)%1
{\color{black}\put(#3,2){\circle*{4}}\put(#1,2){\line(1,0){#2}}}\end{picture}}

@@

\title{How to Learn Useful Changes to Software Projects\\(to Reduce Runtimes and Software Defects)}
\numberofauthors{4} 
\author{  
\alignauthor
Rahul~Krishna\\Tim Menzies, Xipeng~Shen \\
       \affaddr{CS, NC State, USA}\\
       {\{i.m.ralk,~tim.menzies,\\xipengshen\}@gmail.com}
\alignauthor
Andrian Marcus \\
       \affaddr{CS, UT Dallas  } \\ 
       \affaddr{Texas, USA}\\ 
       {amarcus@utdallas.edu}
       \alignauthor
Naveen  \\ Lekkalapudi\\
 \affaddr{Bellhops } \\ 
       \affaddr{Tennessee, USA}\\ 
       {navek91@gmail.com}
\alignauthor
Lucas Layman \\
       \affaddr{Fraunhofer CESE  } \\ 
       \affaddr{College Park, USA}\\ 
       {llayman@fc-md.umd.edu}
\setlength{\columnsep}{7mm}
}
\maketitle
\begin{abstract}
 Business users now demand more insightful
 analytics. Instead of just predicting 
 some result, they also need tools that generate ``plans'';
 i.e. specific suggestions on  what to change  in order to
 improve on the predicted values.

 This paper proposes
 and evaluates the  XTREE planner for software projects. 
XTREE inputs tables of data with independent features plus a   weighted class
that indicate how much each row is  ``bad'' or  ``better''. Plans are edits proposed for each 
row such that the changed row is more likely to be ``better''.  XTREE
learns those plans by building a decision tree across the data, then reporting
the differences in the branches from some current branch to another desired branch.

Using data from 11 software projects,  XTREE found   larger reductions that three
alternate methods.  Those reductions had 
median size of (55\%,38.5\%) and   largest size of (78\%,94\%) in
(defects,runtimes), respectively.
\end{abstract}
\section{Introduction}
Business users are demanding  tools that support    business-level interpretations of their data. At a panel on software analytics at ICSE`12, industrial practitioners lamented the state of the art in software analytics~\cite{menzies12a}. Panelists commented  ``prediction is all well and good, but what about decision making?''. Note that these panelists were more interested in the interpretations and follow-up that occurs after the mining, rather than just the mining itself. So:
\bi
\item
Instead of just accepting {\em predictions} on how many software defects to expect, business users might now demand a {\em plan} to reduce the likelihood of those defects.
\item
Instead of just accepting {\em predictions} on the runtime time of their software, business users might now demand a {\em plan} to reduce that runtime. 
\ei
In response to this business-level demands for planners,
we propose a novel {\em planning} method called XTREE for learning changes to a software system
such that its performance ``improves'', according to some measure. This paper uses XTREE
to reduce  the expected value of the
defects     in Jureczko et al.'s    JAVA systems~\cite{jureczko10};
and the runtimes   in   software    configured by  Siegmund et al.~\cite{sven12}.

The contributions of this paper are (1)  
    the new XTREE    algorithm and (2) an evaluation strategy for   planners.
In the case studies of this paper  our evaluations shows
 XTREE performing significantly ``better'' than    planners  proposed in our prior   work~\cite{me12c,krishna15},
 where ``better'' means 
{\em effective} (plans change the  expected values of the class);
{\em succinct} (it is inconvenient if  plans   always require changing  everything);
and {\em surprising} (a planner should sometimes tell us things we do not expect
since, otherwise,
there is no value added from  using the planner).

The rest of this paper   describes our data, our planners, and the experiment that
ranks XTREE against alternate approaches.  This is followed by notes on related work
and validity.
Note that, to allow for reproducibility, all scripts and data used in this 
study are available on-line at github.com/ai-se/XTREE.

@@

Our planners   use tables of data with independent features and one dependent feature, called the 
``class''. Classes are weighted such that they indicate  what rows are ``bad'' and what rows are ``better''. Plans  change   a row such that it is more likely to be ``better''. 
Specifically, for every test example $Z$,   planners proposes a  plan $\Delta$ to
  adjust   feature $Z_j$:
\[
\forall \delta_j \in \Delta :  Z_j = Z_j + \delta_j
\]
For example, to simplify a  large bug-prone  method, our planners might suggest
to a developer to reduce its size (i.e.  refactor that code by, say, splitting it across
two simpler functions).

@@

This paper is about the next step {\em after} prediction. Suppose
a business user is presented with  a prediction and they do not like what they see; e.g. the runtimes are too long of the number of defects is too high. This user may
then ask a {\em planning}  question; i.e. ``what can we change to do better than that?''.

@@

To answer this question, we note
that data miners explore two  ``clouds'' of data: (1) the cloud of training examples and (2) the  cloud   of test examples
(for a visualization of these clouds, see \fig{howxy}).
\begin{figure}[!t]
  \includegraphics[width=.9\linewidth]{figs/twodee.eps} 
 % \includegraphics[width=1\linewidth]{figs/BDBC.eps}
\caption{Gray, red, green show (1) training examples, (2) test examples and 
  (3) tests that have been altered by planners.
  This figure uses axes generated from the first two components of a PCA analysis of all points. 
}\label{fig:howxy}
\end{figure}
We should mistrust the predictions made by a  model   if it is being applied to examples  that are
too far away from the
training cloud.
To test for ``too far'', we can run a data mining experiment that tests how well
a model learned from the training data applies to the test data. Such experiments return some performance value.

@@

That said, the above comes with some important caveats:
\bi
\item 
We   strongly recommend that predictors are assessed prior to planning. That
issue is explored further in the \tion{tesd}.
\item
Planners should be designed to be {\em trust increasing}. We list four such planning methods in \tion{planners}.
\item
Where possible, planners should be assessed via some external
oracle that can accurately assess new examples. For an example of that kind of analysis,
see  \tion{coc}.
\ei

@@

\caption{Siegmund data.}\label{fig:cpm}
\end{figure}

@@

This section described XTREE (which we call Method4) and  three  
alternate methods for learning plans.
  XTREE  uses the decision tree learner of \fig{where}.D.  It is new to this paper.
  The other methods use the   top-down
	bi-clustering method described in \fig{where}.C  which recursively divides the
	data in two  using a dimension that captures the greatest variability in the data. 
	We proposed Method 1 and 2   in 2012~\cite{me12c} while Methods 3 comes from research conducted earlier this year~\cite{krishna15}.

@@

Method3 connects  each centroid to a nearest neighbor
by a {\em gradient}.
Each gradient has a (bottom,top) end labelled  ($C_i,C_j$) containing the  (worst,best) centroid performance scores, respectively..  
For each test instance, Method3 
find the nearest gradient, the runs up to the top  best end $C_j$, then extracts $B_j$ (which is the
 best-in-cluster associated with  $C_j$).
The returned plan is then computed  from the delta between the test case
and $B_j$.

@@

\newpage \section{Experiments}

@@

\fig{jur} and \fig{conf1} show the effectiveness of our methods seen in 40 repeats with each data set.
In these experiments,   the dependent variables of Jureczko and Siegmund data sets are discrete and continuous in nature, respectively. Hence, while choosing the predictor, we used Random Forest (1) as a classifier for Jureczko data and (2) as a regressor for Siegmund data.
\bi
\item For Siegmund data, we randomized the order of the data, training on one half while identifying treatment plans on the remaining test data. 
\item For the Jureczko data, we used the training and testing sets of \fig{j}. For these datasets,
 all the SMOTE-ing and Random Forest tunings (discussed in \tion{tesd})
occurred in the {\em train} phase of \fig{design}.
\ei
 \subsection{Experimental Results}

@@

@@

\begin{figure*}[!t]
\centering
\includegraphics[width=\linewidth]{figs/Deltas-both.eps}
\caption{Percent frequency for how often certain feature was changed by a plan.}\label{fig:changed}
\end{figure*}

@@

The only coupling issue that XTREE   usually adds to its plans is {\em cbo} (which appears 55\% of the time
in \fig{changed}). But note  that this is {\em object} coupling measure, not class coupling. So here
XTREE is warning against, say, some factory class generating a large
community of agents, all of the same class, who
co-ordinate on some task. This is a different issue to the class redesign issue that would
be triggered by altering {\em loc}.

@@

\subsection{  Evaluation Bias}\label{sect:coc}
Another threat to validity of this work is our use
of predictors learned on the training data to assess the impact of our planners.
This issues was discussed in detail in \tion{trust}.

@@

\subsection{  Feasibility}\label{sect:feasibility}
It is important to also explore the general practicality of the plans. To this end, we are currently undertaking is a \textit{feasibility study} collaboration with industrial partners to assess the general feasibility of our changes. With this study, we would like to understand the nature of changes suggested by XTREE. Specifically: are the changes proposed by XTREE small enough to be easily implemented on software projects? Also, are the changes large enough to be feasible? In other words, we are asking if users can sufficient fine-grained control to make the changes proposed in \fig{changed}.

@@

One problem with these   SBSE methods is that they can  make extensive mutations to the data they are exploring. In the language
of \tion{trust}, these methods may not be {\em trust-increasing} since those algorithms make no attempt
to mutate new examples away from the kinds of data used to commission the model (in which case, we would
start doubting the model's output).

@@

Four planners were assessed here for the tasks of reducing defects and runtimes. 
Three of the methods come from our prior publications~\cite{me12c,krishna15}, and the conclusion of this
paper is that a novel fourth method clearly out-performs the other three
(measured in terms of {\em effectiveness, succinctness, and surprise}).
We conjecture that XTREE worked better than the rest since:
\bi
\item
It uses a supervised method to divide the data and;
\item
When planning how to move examples to better classes, it is   best to use a learner that reflects over 
differences between the classes.
\ei

\section{Questions for Future Work}

XTREE, as used here, sought improvements in a one goal (the   class variable). Does XTREE work for multi-goal reasoning?

 The XTREE algorithm seems quite general to any data of table with rows containing
weighted classes (so we can distinguish ``bad'' rows from ``better'' ones). 
Does   XTREE works  on   domains (other than the  defect/runtime data explored here)?

As an example of a domain that might benefit from XTREE, recent results raise doubts about
the value of changing code to remove ``bad smells''~\cite{Sjoberg13}. Can XTREE be used as a ``really bad smell''
detector to select the subset of possible refactorings that have the most   potential benefit?

As to scalability, XTREE is a post-processor to a decision tree algorithm. Hence, in theory,  XTREE should work on
any domain where data miners can generate decision trees. Given the current state of the art in Big Data,
can   XTREE  be applied to  very large data sets?

There are many more methods for generating plans and
no   one paper can survey them all. For example, this paper has not explored variations to the $\alpha,\beta,\gamma$
parameters that controlled XTREE. Would we get better results if we varied those parameters?

That said, the goal of this paper was not to claim that (e.g.) XTREE is some absolute optimal algorithm. Rather, it is
was to offer a baseline result (with XTREE) and an  evaluation strategy that  can assess  if alternate methods are better than XTREE.
Will other  researchers apply this strategy  to repeat and/or improve
(or even refute) our results?  Perhaps.
